---
title: "Barmah-Millewa Frog Monitoring"
description: |
  Analysis of frog monitoring data from 6 years at Barmah-Millewa forest.
author:
  - name: Justin G Cally
    url: https://justincally.github.io/blog/
    affiliation: Arthur Rylah Institute
    affiliation_url: https://www.ari.vic.gov.au/
date: 2024-05-28
bibliography: frog.bib
output:
  distill::distill_article:
    self_contained: true
    toc: true # table of content true
    toc_float: true # make 
    depth: 3  # upto three depths of headings (specified by #, ## and ###)
    number_sections: false  ## if you want number sections at each table header
    code_folding: true # awesome buttons to show/hide the code
    theme: yeti
slug: bmffrog
knit: |
  (function(input, ...) {
    rmarkdown::render(
      input,
      output_file = 'docs/index.html',
      envir = globalenv()
    )
  })
---

# Brief

This document is intended as supplementary material to the analysis of acoustic frog data between 2018 and 2024 (6 survey years). This document provides relevant R code to process data from the acoustic classifier, extract relevent environmental variables, run a multi-season multi-species occupancy model and interrogate the results.

# Setup

Load relevant R packages and state settings for modeling. 

```{r setup}
library(tidyverse)
library(sf)
library(mapview)
library(terra)
library(readr)
library(readxl)
library(cmdstanr)
library(bayesplot)
library(httr)
library(VicmapR)
library(ggridges)
library(kableExtra)
check_cmdstan_toolchain(fix = TRUE, quiet = TRUE)
options(mc.cores=4)
register_knitr_engine(override = FALSE)
```

Source a range of handy functions for summarily and plotting outputs from a STAN model. These were used in a different project and are stored on github. 

```{r sourcefunctions, results = 'hide', echo =FALSE, message=FALSE}
# Source useful functions from the deer project github page
req <- GET("https://api.github.com/repos/JustinCally/statewide-deer-analysis/git/trees/main?recursive=1")
stop_for_status(req)
filelist <- unlist(lapply(content(req)$tree, "[", "path"), use.names = F)
function_files <- paste0("https://raw.githubusercontent.com/JustinCally/statewide-deer-analysis/main/",
                         grep("functions/", filelist, value = TRUE, fixed = TRUE))
sapply(function_files, source, verbose = F)

source("functions/ppc_plots.R")
```

# Site data  

## Site metadata

Below we load in site metadata for the analysis, which includes the location and deployment dates for the audiomoths. 

```{r SiteData}
site_metadata <- read_excel("data/TLM_Barmah-Millewa_AM_deployment_metadata_all_years.xlsx")

# site_data_habitat <- read_excel("data/site_data/TLM_habitat_transect_data_all_years.xlsx") %>%
#   group_by(Site, Date) %>%
#   summarise(across(c("Depth", "Emergent macrophyte", "Submerged macrophyte",
#                      "Floating macrophyte", "CWD", "Bare ground"), mean)) %>%
#   mutate(Year = as.integer(as.factor(year(Date))))
# 
# site_data_habitat2 <- read_excel("data/site_data/TLM_water_quality_all_years.xlsx") %>%
#   group_by(Site, Date) %>%
#   summarise(across(c("Waterbody width (m)", "Waterbody length (m)", "Max depth (cm)",
#                      "Canopy cover (%)", "Conductivity", "pH", "Water temp (°C)",
#                      "Oxygen (mg/L)", "Oxygen sat (%)", "mmHg", "Turbidity (FNU)",
#                      "Turbidity (NTU)"), mean)) %>%
#   mutate(Year = as.integer(as.factor(year(Date))))

# Read in water management area data 
WMAs <- read_excel("data/WMAs/Sites_WMAs.xlsx")
WMA_scores <- read_excel("data/WMAs/WMA_flood_scores.xlsx")

WMA_joined <- left_join(WMAs, WMA_scores) %>% 
  select(WMA, `2018-19` = `2018`, `2019-20` = `2019`, `2020-21` = `2020`, `2021-22` = `2021`, `2022-23` = `2022`, `2023-24` = `2023`) %>% 
  pivot_longer(2:7, names_to = "Monitoring year", values_to = "FloodScore") %>%
  distinct()

# Convert site metadata to a spatially explicit dataset and format
site_metadata_sf <- site_metadata %>%
  left_join(WMA_joined, by = join_by(WMA, `Monitoring year`)) %>% 
  st_as_sf(coords = c("X", "Y"), crs = 28355) %>%
  mutate(`Site name` = case_when(`Site` == "Warri" ~ "Warri Yards", 
                                 `Site` == "Hughes Crossing" ~ "Hughs Crossing",
                                 Site == "Little Rushy Swamp" ~ "Little Rushy",
                                 Site == "Reed Beds Swamp" ~ "Reed Bed Swamp",
                                 Site == "Wathours Lagoon" ~ "Wathours",
                                 TRUE ~ `Site`), 
         Site = paste(`Site name`, AM, sep = "_AM"), 
         Site_Year = paste(Site, `Monitoring year`, sep = "_"),
         Year = factor(`Monitoring year`) %>% as.integer(),
         Cluster = factor(`Site name`) %>% as.integer(),
         Loc = factor(Site) %>% as.integer(),
         No_survey_nights = as.numeric(No_PAM_nts)) %>%
  # arrange(`Site`) %>%
  filter(!is.na(No_survey_nights) & No_survey_nights > 0) %>%
  arrange(Site_Year) %>%
  filter(!(Site_Year %in% c("Rookery_AM1_2022-23", "Two Mile Bunyip_AM2_2021-22"))) 

unique_sites <- site_metadata_sf %>% 
  group_by(Site) %>%
  slice(1) %>%
  select(Site)

# get alist of the nights deployed 
nights_dep <- list()
for(i in 1:nrow(site_metadata_sf)) {
  nights_dep[[i]] <- data.frame(Site = site_metadata_sf$Site[i], 
                                Site_Year = site_metadata_sf$Site_Year[i], 
                           Date = seq.Date(from = as.Date(site_metadata_sf$`Date start`[i]), 
                                           by = "1 day", 
                                           length.out = site_metadata_sf$No_survey_nights[i])) #site_metadata_sf$No_survey_nights[i]))
}
nights_dep_joined <- bind_rows(nights_dep)
```

## Generate a hull for the study area

```{r hull, eval = F}
site_hull <- st_convex_hull(unique_sites %>% st_combine()) %>% 
  st_transform(3577) %>%
  st_buffer(100)
hydro_data <- vicmap_query("open-data-platform:hy_water_area_polygon") %>%
  filter(BBOX(site_hull)) %>%
  collect()

elev_data <- vicmap_query("open-data-platform:el_contour") %>%
  filter(BBOX(site_hull)) %>%
  collect()

mapview(hydro_data, zcol = "feature_type_code") + mapview(site_metadata_sf)

sf::st_write(site_hull, "data/barmah_hull/barmah_hull.shp")
sf::st_write(site_hull %>% st_transform(4326), "data/barmah_hull/barmah_hull.geojson")
sf::st_write(unique_sites %>% st_transform(3577) %>% st_buffer(100), "data/site_metadata_sf/site_metadata_sf.shp", overwrite = T)

sf::st_write(unique_sites %>% st_buffer(10, endCapStyle = "SQUARE") %>% st_transform(4326) %>% st_combine(), "data/barmah_hull/site_locs.geojson", overwrite = T)
```

## Hydrological data 

Read in the hydrological data. This data has been obtained through Digital Earth Austraia sandbox (https://app.sandbox.dea.ga.gov.au/), the code used to extract the hydrologial data and the csvs for each site are available on GitHub: https://github.com/JustinCally/dea-barmah  

```{r hydro, message = F, warning = F, eval = F}
# load in csv files of hydro conditions 
# list.files("https://github.com/JustinCally/dea-barmah/tree/main/output")
# req <- GET("https://api.github.com/repos/JustinCally/dea-barmah/git/trees/main?recursive=1")
# stop_for_status(req)
# filelist <- unlist(lapply(content(req)$tree, "[", "path"), use.names = F)
# csv_files <- paste0("https://raw.githubusercontent.com/JustinCally/dea-barmah/main/",
#                          grep("output/", filelist, value = TRUE, fixed = TRUE))
# csv_files <- grep(".ipynb_checkpoints", x = csv)
# 
# sapply(function_files, source, verbose = F)

#paste
csv_files <- paste0("https://raw.githubusercontent.com/JustinCally/dea-barmah/main/output/", unique(site_metadata_sf$Site), ".csv")
csv_files <- stringr::str_replace_all(csv_files, pattern = " ", replacement = "%20")
hydryo_data <- list()
for(i in 1:length(csv_files)) {
  hydryo_data[[i]] <- readr::read_csv(csv_files[i], show_col_types = F) %>%
    mutate(Site = unique(site_metadata_sf$Site)[i])
}

hydro_data_combined <- bind_rows(hydryo_data)
```

Given that the data was often only captured every 2 weeks we want to obtain daily estimates of hydrological conditions. To this we interpolate the hydrological data across the time range of the study (2018 - early 2024). We do this by fitting a polynomial regression with the `loess()` function. 

```{r interpolation, eval = F}
#hydro interpolation 

new_df_raw <- data.frame(date = seq(from = min(hydro_data_combined$date),
                                to = max(hydro_data_combined$date),
                                by="day"))

interpolate_data <- list()

for(i in 1:length(csv_files)) {
  
  hd <- hydryo_data[[i]]
  
  mod_pv <- loess(pv ~ as.numeric(date), data = hd, span = 0.075)
  mod_npv <- loess(npv ~ as.numeric(date), data = hd, span = 0.075)
  mod_bs <- loess(bs ~ as.numeric(date), data = hd, span = 0.075)
  mod_wet <- loess(wet ~ as.numeric(date), data = hd, span = 0.075)
  mod_water <- loess(water ~ as.numeric(date), data = hd, span = 0.075)
  mod_veg_areas <- loess(veg_areas ~ as.numeric(date), data = hd, span = 0.075)


interpolate_data[[i]] <- new_df_raw %>%
  mutate(Site = unique(site_metadata_sf$Site)[i],
         pv = predict(mod_pv, newdata = new_df_raw),
         npv = predict(mod_npv, newdata = new_df_raw),
         bs = predict(mod_bs, newdata = new_df_raw),
         wet = predict(mod_wet, newdata = new_df_raw),
         water = predict(mod_water, newdata = new_df_raw),
         veg_areas = predict(mod_veg_areas, newdata = new_df_raw)) %>%
  mutate(across(.cols = c(pv,npv,bs, wet, water, veg_areas), 
                .fns = ~ case_when(.x > 1 ~ 1, 
                                   .x < 0 ~ 0,
                                   TRUE ~ .x)))

}

interpolate_combined <- bind_rows(interpolate_data) %>%
  mutate(Date = as.Date(date)) %>%
  select(-date)

interpolate_data[[62]] %>%
  ggplot() +
  geom_point(aes(x = date, y = pv)) +
  ylim(c(0,1))
```

## Daily climate variables  

Using the National Oceanic and Atmospheric Administration (NOAA) data of gridded daily precipitation and temperature, we extract the daily precipitation and minimum temperature values estimated at each site at a broad spatial scale (0.5 x 0.5 degrees).

```{r precip, eval = F}
# load in daily precip 
# https://psl.noaa.gov/data/gridded/data.cpc.globalprecip.html
precipitation_daily <- c(terra::rast("data/climate/precip_2018.nc"), 
                         terra::rast("data/climate/precip_2019.nc"), 
                         terra::rast("data/climate/precip_2020.nc"),
                         terra::rast("data/climate/precip_2021.nc"), 
                         terra::rast("data/climate/precip_2022.nc"), 
                         terra::rast("data/climate/precip_2023.nc"),
                         terra::rast("data/climate/precip_2024.nc"))
precipitation_daily_df <- bind_cols(site_metadata_sf$Site_Year, 
                                    extract(precipitation_daily, vect(site_metadata_sf)))
colnames(precipitation_daily_df) <- c("Site_Year","ID", as.character(seq.Date(from = as.Date("2018-01-01"), 
                                                     to = as.Date("2024-05-13"), by = "day")))

precipitation_daily_long <- pivot_longer(precipitation_daily_df, 
                                         cols = 3:2327, names_to = "Date", values_to = "Precipitation") %>%
  select(-ID) %>%
  mutate(Date = as.Date(Date)) %>%
  right_join(nights_dep_joined, by = join_by(Site_Year, Date))
```

```{r temp, eval = F}
# load in daily precip 
# https://psl.noaa.gov/data/gridded/data.cpc.globalprecip.html
tmin_daily <- c(terra::rast("data/climate/precip_2018.nc"), 
                         terra::rast("data/climate/tmin_2019.nc"), 
                         terra::rast("data/climate/tmin_2020.nc"),
                         terra::rast("data/climate/tmin_2021.nc"), 
                         terra::rast("data/climate/tmin_2022.nc"), 
                         terra::rast("data/climate/tmin_2023.nc"),
                         terra::rast("data/climate/tmin_2024.nc"))
tmin_daily_df <- bind_cols(site_metadata_sf$Site_Year, 
                                    extract(tmin_daily, vect(site_metadata_sf)))
colnames(tmin_daily_df) <- c("Site_Year","ID", as.character(seq.Date(from = as.Date("2018-01-01"), 
                                                     to = as.Date("2024-05-13"), by = "day")))

tmin_daily_long <- pivot_longer(tmin_daily_df, 
                                cols = 3:2327, 
                                names_to = "Date", 
                                values_to = "tmin") %>%
  select(-ID) %>%
  mutate(Date = as.Date(Date)) %>%
  right_join(nights_dep_joined, by = join_by(Site_Year, Date))
```

## Flow Data 

We also include two key flow variables in our analysis: 
1. Whether the site was subject to any active regulation  
2. When that active regulation occured via release of water through regulators. 

```{r flow}
flow_hist_b <- read_excel("data/reg_data/ALL_REG_DATA.xlsx", sheet = 1)
flow_hist_m <- read_excel("data/reg_data/ALL_REG_DATA.xlsx", sheet = 2)
flow_hist <- full_join(flow_hist_b, flow_hist_m) %>%
  filter(Date > as.Date("2018-01-01")) %>%
  mutate(Days = Date + lubridate::days(2))

reg_site_sizes <- read_excel("data/reg_data/Reg_Site Interactions.xlsx") %>%
  mutate(gate_1_total = `No gates Size 1` * `Gate size 1 height` * `Gate size 1 width`, 
         gate_2_total = `No gates Size 2` * `Gate size 2 height` * `Gate size 2 width`,
         gate_3_total = `No gates Size 3` * `Gates Size 3 height` * `Gates Size 3 width`,
         gate_4_total = `No gates Size 4` * `Gates Size 4 height` * `Gates Size 4 width`,
         gate_5_total = `No gates Size 5` * `Gates Size 5 height` * `Gates Size 5 width`, 
         gate_avg = sum(gate_1_total,gate_2_total,gate_3_total,gate_4_total,gate_5_total, na.rm = T)/`Total gates`, 
         gate_avg = coalesce(gate_avg, 0)) %>%
          select(regulator = `Reg and overbank site details`, gate_avg)

reg_to_site <- read_excel("data/reg_data/Reg_Site Interactions.xlsx", sheet = 2) %>%
  filter(`Regulation score` == "3")
reg_to_site_long <- reg_to_site %>%
  pivot_longer(cols = 9:47, names_to = "regulator", values_to = "count", values_drop_na = T) %>%
  select(Site = SITE, water_arrival = `Water arrival (days)`, regulator, count) %>%
  mutate(`Site` = case_when(`Site` == "Warri" ~ "Warri Yards", 
                                 `Site` == "Hughes Crossing" ~ "Hughs Crossing",
                                 Site == "Little Rushy Swamp" ~ "Little Rushy",
                                 Site == "Reed Beds Swamp" ~ "Reed Bed Swamp",
                                 Site == "Wathours Lagoon" ~ "Wathours",
                                 TRUE ~ `Site`)) %>%
  full_join(reg_site_sizes)

flow_hist_long <- flow_hist %>%
  pivot_longer(cols = 5:34, names_to = "regulator", values_to = "n_gates") %>%
  left_join(reg_to_site_long) %>%
  group_by(Date, Site) %>% 
  summarise(total_gate_size_on = sum(n_gates * gate_avg)) %>%
  ungroup()

flow_hist_long_c <- bind_rows(flow_hist_long %>% mutate(Site = paste0(Site, "_AM1")), 
                              flow_hist_long %>% mutate(Site = paste0(Site, "_AM2"))) %>%
  mutate(Regulated = TRUE)
```

## Save daily variables 

Given some of the preceding code can take a while to run we save the output as an rds. This output is the daily values at each site for hydrological and climate conditions as well as the days since the start of spring (1st of September). 

```{r dailyvars, eval = F}
daily_vars <- left_join(precipitation_daily_long, 
                        tmin_daily_long) %>%
  mutate(Day = lubridate::yday(Date)-243, 
         Day = case_when(Day < 1 ~ Day + 243 + (365-243), 
                         TRUE ~ Day),
         cosDay = 2*pi*Day/365, 
         sinDay = 2*pi*Day/365) %>% 
  inner_join(interpolate_combined, by = c("Site", "Date")) %>%
  distinct()

saveRDS(daily_vars, "data/intermediate/daily_vars.rds")
```

## Load daily variables

```{r loaddailyvars}
# read in daily vars
daily_vars <- readRDS("data/intermediate/daily_vars.rds") %>%
  left_join(flow_hist_long_c) %>%
  mutate(total_gate_size_on = coalesce(total_gate_size_on, 0), 
         any_gates_on = total_gate_size_on > 0, 
         Regulated = coalesce(Regulated, FALSE))

site_vars <- daily_vars %>%
  group_by(Site_Year) %>%
  summarise(across(c("pv", "npv", "bs", "wet", "water", "veg_areas"), mean)) %>% 
  ungroup() %>%
  mutate(watercomb = wet + water) %>%
  left_join(site_metadata_sf %>% 
              st_drop_geometry() %>% 
              transmute(Site_Year, WMA, FloodScore = factor(FloodScore), 
                        Site_type = factor(Site_type), Water_permanence = factor(Water_permanence)), 
            by = join_by(Site_Year))
```

# Acoustic data 

## Daily acoustic data

For this analysis we model occupancy and call rate using daily maximum call probability scores. We obtained these scores from the raw model output and processed them in a separate script available in this repository: `frog_daily_subset.R`. This data comes in two chunks (2018-2021 and 2021-2024). 

```{r dailymaxscores, warning = F, message = F}
# Subset to site years
site_year <- nights_dep_joined %>% select(Site, Site_Year, Date) %>% 
  unique() %>% 
  left_join(site_metadata_sf %>% st_drop_geometry() %>% select(Site_Year, Year, Cluster, Loc), by = join_by(Site_Year))

# List of species we have data for

species_list <- c(`Barking Marsh Frog` = 13059,
                  `Eastern Sign-bearing Froglet` = 13131,
                  `Common Froglet` = 13134,
                  # `Sloane's Froglet` = 13135,
                  `Common Spadefoot Toad` = 13086,
                  `Peron's Tree Frog` = 13204,
                  `Pobblebonk Frog` = 63913,
                  `Spotted Marsh Frog (race unknown)` = 13063)

nspecies <- length(species_list)

sp_l_col <- paste0("sp_", species_list)

# species_cols <- str_extract_all(last(colnames(site_records)), "[-+]?[0-9]+")[[1]] 

site_records_all <- list()
daily_files <- list.files("data/daily_data/OtherYears/", full.names = T)
files_to_read <- list.files("data/daily_data/OtherYears/")
sp_read_order <- stringr::str_remove_all(files_to_read, c("daily_max_|.csv"))

for(i in 1:length(daily_files)) {
  site_records_all[[i]] <- read_csv(daily_files[i], show_col_types = F) %>% 
    select(-1) %>% 
    mutate(DateTime = as.POSIXct(DateTime, tz = Sys.timezone()), 
           Date = coalesce(Date, as.Date(DateTime - hours(12)))) %>%
    mutate(`Site` = case_when(Site == "Little Rushy Swamp_AM1" ~ "Little Rushy_AM1",
                              Site == "Little Rushy Swamp_AM2" ~ "Little Rushy_AM2",
                              TRUE ~ `Site`)) %>%
    mutate(Validated_Grp = as.character(species_list[[sp_read_order[i]]]),
           Validation_Species = sp_read_order[i],
           Validated_Grp_score = !!sym(sp_read_order[i])) %>%
    select(FileId, Filename, Site, Date, DateTime, Orig_Start, Orig_End, 
           Validated_Grp, Validation_Species, Validated_Grp_score) %>% 
    left_join(site_year) %>%
    filter(!is.na(Year))
}

names(site_records_all) <- sp_read_order
 
#### 2021-24 ####
site_records_2023 <- list()
daily_files <- list.files("data/daily_data/2021-2024/", full.names = T)
files_to_read <- list.files("data/daily_data/2021-2024/")
sp_read_order <- stringr::str_remove_all(files_to_read, c("daily_max_|.csv"))

for(i in 1:length(daily_files)) {
  site_records_2023[[i]] <- read_csv(daily_files[i], show_col_types = F) %>% 
    select(-1) %>% 
    mutate(Validated_Grp = as.character(species_list[[sp_read_order[i]]]),
           Validation_Species = sp_read_order[i],
           Validated_Grp_score = !!sym(sp_read_order[i])) %>% 
    select(FileId, Filename, Site, Date, DateTime, Orig_Start, Orig_End, 
           Validated_Grp, Validation_Species, Validated_Grp_score) %>% 
    left_join(site_year) %>%
    filter(!is.na(Year))
}

names(site_records_2023) <- sp_read_order

site_records_combined <- list()

for(x in names(species_list)) {
  site_records_combined[[x]] <- bind_rows(site_records_all[[x]], site_records_2023[[x]]) %>%
    mutate(SnippetID = paste(FileId, Site, Date, Orig_Start, Orig_End, sep = "_"))
}



```

# Validated site records  

Up to 500 daily calls for each species were validated, with validations for each species covering ranges of probability scores from 0 to 1. We read in these validations and visualise the distribution for 'true positives' and 'false positives'.

```{r validatedscores, fig.width = 8, fig.height = 5, fig.cap = "Distributions of probability scores assigned by the classifier"}
# site_validation <- read_excel("data/TLM_Barmah-Millewa_2022-23_validations_ALL.xlsx") %>%
#  mutate(Site = case_when(Site == "Warri" ~ "Warri Yards", 
#                          TRUE ~ Site),
#         Site = paste(Site, AM, sep = "_AM"))

files_to_read <- list.files("data/validated_calls/OtherYears")
sp_read_order <- stringr::str_remove(files_to_read, "_validation.csv")

validated_calls <- list()
for(i in 1:nspecies) {
  validated_calls[[sp_read_order[i]]] <- readr::read_csv(paste0("data/validated_calls/OtherYears/", files_to_read[i]), 
                                                         show_col_types = F) %>%
    mutate(Validated_Grp = as.character(species_list[[sp_read_order[i]]]),
           Validation_Species = sp_read_order[i],
           Validated_Grp_score = !!sym(sp_read_order[i]), 
           Date = as.Date(Date, "%d/%m/%Y")) %>%
    mutate(Detected = as.integer(stringr::str_detect(Validation_TaxonIDs, Validated_Grp)), 
           SnippetID = paste(FileId, Site, Date, Orig_Start = Orig_Start_Time, Orig_End = Orig_End_Time, sep = "_")) %>%
    select(Validated_Grp, Validation_Species, Validated_Grp_score, SnippetID, Detected) 
    
}



files_to_read_2 <- list.files("data/validated_calls/2021-2024")
sp_read_order_2 <- stringr::str_remove(files_to_read_2, "_validation.csv")

validated_calls_2 <- list()
for(i in 1:nspecies) {
  validated_calls_2[[sp_read_order_2[i]]] <- validated_calls[[sp_read_order_2[i]]] %>% bind_rows(readr::read_csv(paste0("data/validated_calls/2021-2024/", files_to_read_2[i]), 
                                                         show_col_types = F) %>%
    mutate(Validated_Grp = as.character(species_list[[sp_read_order_2[i]]]),
           Validation_Species = sp_read_order_2[i],
           Validated_Grp_score = !!sym(sp_read_order_2[i]), 
           Date = as.Date(Date, "%d/%m/%Y")) %>%
    mutate(Detected = as.integer(stringr::str_detect(Validation_TaxonIDs, Validated_Grp)), 
           SnippetID = paste(FileId, Site, Date, Orig_Start = Orig_Start_Time, Orig_End = Orig_End_Time, sep = "_")) %>%
    select(Validated_Grp, Validation_Species, Validated_Grp_score, SnippetID, Detected))
    
}

records_validated <- list()

for(x in names(species_list)) {
  records_validated[[x]] <- left_join(site_records_combined[[x]], validated_calls_2[[x]]) %>%
    mutate(Detected = coalesce(Detected, 2L)) %>%
    arrange(Site_Year)
}

validated_calls_combined <- bind_rows(records_validated) %>%
  filter(Detected != 2) %>%
  rowwise() %>%
  mutate(Prob_f = case_when(Validated_Grp_score == 1 ~ 0.999,
                            Validated_Grp_score == 0 ~ 0.001,
                     TRUE ~ Validated_Grp_score),
        Prob_One_M = qlogis(Prob_f)) 

validated_calls_combined %>%
  mutate(Detected = as.factor(Detected)) %>%
  filter(Validation_Species != "Common Spadefoot Toad") %>%
  ggplot() +
  geom_density(aes(x = Prob_f, fill = Detected, colour = Detected), position = "jitter", alpha = 0.25) +
  facet_wrap(~Validation_Species) +
  xlab("Classifier-assigned probability") +
  delwp_theme() 

```

# Format data  

For the model, we filter out common spadefoot toad as we have no validated detections for this species. Below we format the acoustic, site, and nightly data into a format that can be used in the custom STAN model. 

```{r formatdata, message = F, warning = F}
# Generate STAN data 
sp_out <- c("Common Spadefoot Toad") 
            # "Pobblebonk Frog", 
            # "Spotted Marsh Frog (race unknown)")
species_list_cleaned <- species_list[which(!(names(species_list) %in% sp_out))]
records_validated <- records_validated[which(!(names(records_validated) %in% sp_out))]

nspecies_cleaned <- length(species_list_cleaned)

sp_l_col_cleaned <- paste0("sp_", species_list_cleaned)

# joined data
joined_data <- list()
site_data <- list()

# indexes for species
site_names <- unique(records_validated[[1]]$Site_Year)
nsites <- length(unique(records_validated[[1]]$Site_Year))

empty_array <- array(dim = c(nsites, length(species_list_cleaned)))
start_idx_0 <- empty_array
end_idx_0 <- empty_array
start_idx_1 <- empty_array
end_idx_1 <- empty_array
start_idx_2 <- empty_array
end_idx_2 <- empty_array
any_seen <- empty_array
site_idx_start <- empty_array
site_idx_end <- empty_array


call_rate_matrix <- list()
occ_matrix <- list()

joined_data <- readRDS("data/intermediate/joined_data.rds")

for(i in 1:length(species_list_cleaned)) {

### Because we are randomly sampling ####
# joined_data[[i]] <- records_validated[[i]] %>%
#   inner_join(daily_vars %>% na.omit()) %>%
#   mutate(Prob_One = Validated_Grp_score,
#          Site_f = as.integer(factor(Site_Year)),
#          Prob_f = case_when(Prob_One == 1 ~ 0.999,
#                             Prob_One == 0 ~ 0.001,
#                             TRUE ~ Prob_One),
#          Prob_One_M = qlogis(Prob_f),
#          weight = case_when(Detected == 2 ~ 0.01,
#                             TRUE ~ 1)) %>%
#   group_by(Site_f) %>%
#   mutate(any_seen = case_when(any(Detected == 1) ~ 1,
#                               TRUE ~ 0)) %>%
#   slice_sample(n = 30, weight_by = weight) %>%
#   ungroup() %>%
#   arrange(Site_f, Detected)

site_data[[i]] <- joined_data[[i]] %>%
  mutate(Loc = as.factor(Site) %>% as.integer(), 
         Cluster = as.factor(str_remove_all(Site, "_AM1|_AM2")) %>% as.integer()) %>%
  select(Site_f, Loc, Cluster, Year, Site_Year, any_seen, Site) %>%
  distinct() %>%
  arrange(Site_f) %>% 
  left_join(site_vars, by = "Site_Year")

# occupancy model
psi_form <- ~ scale(sqrt(wet)) + scale(pv) + scale(sqrt(water)) + Site_type + Water_permanence  # + FloodScore
occ_matrix[[i]] <- model.matrix(psi_form, data = site_data[[i]])

# Daily precipitation matrix
theta_form <- ~ scale(sqrt(Precipitation)) + scale(tmin) + scale(sqrt(wet)) + scale(sqrt(water)) + scale(pv) + Regulated + any_gates_on + Regulated*any_gates_on #+ cosDay + sinDay
call_rate_matrix[[i]] <- model.matrix(theta_form, data = joined_data[[i]])

for(j in 1:nsites) {
  start_idx_0[j,i] <- min(which(joined_data[[i]]$Site_f == j & joined_data[[i]]$Detected == 0))
  end_idx_0[j,i] <- max(which(joined_data[[i]]$Site_f == j & joined_data[[i]]$Detected == 0))
  
  start_idx_1[j,i] <- min(which(joined_data[[i]]$Site_f == j & joined_data[[i]]$Detected == 1))
  end_idx_1[j,i] <- max(which(joined_data[[i]]$Site_f == j & joined_data[[i]]$Detected == 1))
  
  start_idx_2[j,i] <- min(which(joined_data[[i]]$Site_f == j & joined_data[[i]]$Detected == 2))
  end_idx_2[j,i] <- max(which(joined_data[[i]]$Site_f == j & joined_data[[i]]$Detected == 2))
    
  site_idx_start[j,i] <- min(which(joined_data[[i]]$Site_f == j))
  site_idx_end[j,i] <- max(which(joined_data[[i]]$Site_f == j))
}

start_idx_0[is.infinite(start_idx_0[,i]),i] <- 0
start_idx_1[is.infinite(start_idx_1[,i]),i] <- 0
start_idx_2[is.infinite(start_idx_2[,i]),i] <- 0

end_idx_0[is.infinite(end_idx_0[,i]),i] <- 0
end_idx_1[is.infinite(end_idx_1[,i]),i] <- 0
end_idx_2[is.infinite(end_idx_2[,i]),i] <- 0

any_seen[,i] <- as.integer(start_idx_1[,i] != 0)
}

# saveRDS(joined_data, "data/intermediate/joined_data.rds")
```

In addition to the validation data, we also have previous validations at a site-level that notes whether any calls were detected at that site in that year. These are included in the model as if restricts the probability space we need to marginalize over for sites that have at least one detection. 

```{r anyseen}
# read in updated any seen validations 
any_seen_add <- readr::read_csv("data/any_seen/sp_detect_per_site_per_year.csv") %>% 
      mutate(`Site_AM` = case_when(Site_AM == "Little Rushy Swamp_AM1" ~ "Little Rushy_AM1",
                              Site_AM == "Little Rushy Swamp_AM2" ~ "Little Rushy_AM2",
                              Site_AM ==  "Hughes Crossing_AM1" ~ "Hughs Crossing_AM1",
                              Site_AM ==  "Hughes Crossing_AM2" ~ "Hughs Crossing_AM2",
                              Site_AM == "Reed Beds Swamp_AM1" ~ "Reed Bed Swamp_AM1",
                              Site_AM == "Reed Beds Swamp_AM2" ~ "Reed Bed Swamp_AM2",
                              Site_AM == "Wathours Lagoon_AM1" ~ "Wathours_AM1",
                              Site_AM == "Wathours Lagoon_AM2" ~ "Wathours_AM2",
                              TRUE ~ `Site_AM`)) %>%
  mutate(Site_Year = paste(Site_AM, Year, sep = "_")) %>%
  arrange(Site_Year) 

setdiff(site_names, any_seen_add$Site_Year) %>% paste(collapse = "\n") %>% cat()

any_seen_clean <- any_seen_add %>% filter(Site_Year %in% site_names) %>%
  select(`Barking Marsh Frog`,
         `Eastern Sign-bearing Froglet` = `Eastern Sign-Bearing Froglet`,
         `Common Froglet`, 
         `Peron's Tree Frog`, 
         `Pobblebonk Frog` = Pobblebonk,
         `Spotted Marsh Frog (race unknown)` = `Spotted Marsh Frog`) %>%
  as.matrix()

any_seen_joined <- any_seen_clean

for(j in 1:ncol(any_seen_clean)) {
  for(i in 1:nrow(any_seen_clean))
    any_seen_joined[i,j] <- max(any_seen[i,j], any_seen_clean[i,j])
}
```

We then combine all the data into a list, which will be used by the STAN model 

```{r stanlist}
nfiles <- nrow(joined_data[[1]])
scores <- lapply(joined_data, function(x) pull(x, Prob_One_M)) 
site_list <- lapply(joined_data, function(x) pull(x, Site_f)) 
loc_code <- lapply(site_data, function(x) pull(x, Loc)) 
cluster_code <- lapply(site_data, function(x) pull(x, Cluster)) 
year_code <- lapply(site_data, function(x) pull(x, Year)) 
detected <- lapply(joined_data, function(x) pull(x, Detected))
site_code <- lapply(joined_data, function(x) pull(x, Site_f))

nyear <- length(unique(unlist(year_code)))
nloc <- length(unique(unlist(loc_code)))
nclusters <- length(unique(unlist(cluster_code)))


stan_data <- list(nfiles = nfiles, 
                  nsites = nsites, 
                  nspec = nspecies_cleaned,
                  nyear = nyear, 
                  nloc = nloc, 
                  nclusters = nclusters,
                  score = array(unlist(scores),dim=c(nfiles,nspecies_cleaned)), 
                  site_code = array(unlist(site_code),dim=c(nfiles,nspecies_cleaned)),
                  loc_code = array(unlist(loc_code),dim=c(nsites,nspecies_cleaned)), 
                  cluster_code = array(unlist(cluster_code),dim=c(nsites,nspecies_cleaned)), 
                  year_code = array(unlist(year_code),dim=c(nsites,nspecies_cleaned)), 
                  year_counts = unname(apply(array(unlist(year_code),dim=c(nsites,nspecies_cleaned)),2,table)),
                  year_counts_file = unname(as.integer(table(joined_data[[1]]$Year))),
                  site_idx_start = site_idx_start, 
                  site_idx_end = site_idx_end,
                  start_idx_0 = start_idx_0, 
                  end_idx_0 = end_idx_0,
                  start_idx_1 = start_idx_1, 
                  end_idx_1 = end_idx_1, 
                  start_idx_2 = start_idx_2, 
                  end_idx_2 = end_idx_2, 
                  any_seen = any_seen_joined, 
                  theta_mm = call_rate_matrix, 
                  theta_nc = ncol(call_rate_matrix[[1]]), 
                  psi_mm = occ_matrix, 
                  psi_nc = ncol(occ_matrix[[1]]), 
                  detected = array(unlist(detected),dim=c(nfiles,nspecies_cleaned)))

# stan_data$year_code[stan_data$year_code == 5] <- 4
stan_data_beta <- stan_data 
stan_data_beta$score <- array(unlist(lapply(joined_data, function(x) pull(x, Prob_f))),dim=c(nfiles,nspecies_cleaned))

saveRDS(stan_data, "data/stan_data.rds")

```

# STAN Model  

Below is the STAN model used in our analysis. It is a multi-species, multi-season model. We model observations as a mixture of validated detections and unvalidated probable detections, using the probability scores assigned by the acoustic classifier as the response variable in a beta regression 

```{cmdstan, file = "stan/continuous_model_ms_my_beta.stan", output.var = "cont_model_ms_beta", eval = FALSE}

```

```{r readmods, echo = F}
# cont_model_ms_norm <- cmdstan_model("stan/continuous_model_ms_my_norm.stan")
cont_model_ms_beta <- cmdstan_model("stan/continuous_model_ms_my_beta.stan")
# cont_model_ms_normal <- cmdstan_model("stan/continuous_model_ms_my_normal.stan")
```

```{r modelparams}
ni <- 250 #samples
nw <- 250 #warmups
nc <- 8 #chains
```

```{r fitmods, eval = F}
# fit_normal <- cont_model_ms_normal$sample(data = stan_data_beta,
#                          chains = nc,
#                          parallel_chains = nc,
#                          show_messages = TRUE,
#                          save_warmup = FALSE,
#                          iter_sampling = ni,
#                          iter_warmup = nw, refresh = 50)
# 
# fit_normal$save_object("outputs/fit_normal.rds")

fit_beta <- cont_model_ms_beta$sample(data = stan_data_beta,
                         chains = nc,
                         parallel_chains = nc,
                         show_messages = TRUE,
                         save_warmup = FALSE,
                         iter_sampling = ni,
                         iter_warmup = nw, refresh = 50)

fit_beta$save_object("outputs/fit_beta.rds")
```

```{r readinmods}
fit_beta <- readRDS("outputs/fit_beta.rds")
```

# Analysis summary. 

## Data 

### Acoustic surveys  

Six consecutive years of acoustic data was analysed. Acoustic monitoring began in the Austral Spring/Summer of 2018-2019 and has continued through until the most recent 2023-2024 season. Throughout the six-year monitoring period a total of 36 sites with paired songmeters (n = 72) were surveyed. However, not each songmeter-site was surveyed each year, or in some cases equipment errors gave no survey data, leading to a total of 234 songmeter-sites being surveyed across the six years. Songmeters were deployed for varying lengths of time, with an average of 124 nights and a maximum of 241 nights. Acoustic recordings were processed using an automated species-classification recogniser. 

The output of this acoustic classifier model produced 11,633,143 snippets of possible frog detections for the following species: Barking Marsh Frog, Eastern Sign-bearing Froglet, Common Froglet, Common Spadefoot Toad, Peron's Tree Frog, Pobblebonk Frog, Spotted Marsh Frog (race unknown). Each snippet was encoded with a vector of 'probability scores' for each species; which describe the probabilistic assignment of a given call snippet to a given species. For this analysis, we subset the over 11-million records into a snippet with the species-level daily maximum probability score, leaving us with 29,213 possible call snippets for each species across all site-year combinations. For each species between 338 and 441 call snippets were selected using stratified random sampling to be validated. To ensure we were validating a range of 'probability scores' for each species this stratification enabled balanced sampling across five 'probability score' strata (0 - 0.2, 0.2 - 0.4, 0.4 - 0.6, 0.6 - 0.8 and 0.8 - 1). 

The final data includes a mix of validated and not validated snippets for each species and for each songmeter-site across the surveys years. Each snippet has a 'probability score' between 0 - 1 as well as a validation label (0 = validated: species not heard, 1 = validated: species heard, and 2 = not validated). 

### Site data 

We collected and compiled data that may vary spatially and temporally. Data compiled can be used in the modelling of either; (i) occupancy or (ii) call rate/activity. Whereby data associated with occupancy will be related to conditions during the span of the survey in that year (e.g. multiple months). Alternatively data associated with call rate was collected at a daily/nightly frequency, as it seeks to explain variation in calling activity over the length of deployment within an occupied site. 

#### Hydrological data  

Hydrology is expected to be a key driver of frog occupancy and activity within Barmah-Millewa. In this analysis we used remotely-sensed estimates of water, wetness and photosynthetic vegetation [@geoscience]. This process was supported by the use of Digital Earth Australia's Wetlands Insight Tool (WIT). Using Landsat 5,7, and 8, we estimated fractional cover of freestanding water, wetness, photosynthetic vegetation, dry vegetation and bare soil. For each songmeter-site we obtained time-series data across the six survey years within a 100m buffer surrounding the songmeter. Landsat measurements occurred approximately every 15 days. To estimate the fractional cover metrics on a daily-level we interpolated the data between measurements. We implemented local polynomial regression to enable daily predictions of fractional vegetation and water coverage using the `loess()` R function [@Cleveland2017].      

#### Active regulation data 

Active regulation data was obtained via a time-series of regulator usage and data linking each regulator with sites that would be impacted by their usage. Active regulation data as available on a daily-basis, with a uniform 2-day lag implemented between the regulator function (ON/OFF) and any modelled effect on sites. 

#### Daily climate data 

Daily temperature and precipitation are anticipated to impact frog calling activity. To obtain estimates of daily minimum temperature and precipitation we used National Oceanic and Atmospheric Administration (NOAA) data of gridded daily precipitation and temperature [@noaa]. Daily precipitation and minimum temperature values were estimated at each site at a broad spatial scale (0.5 x 0.5 degrees), thus the variation in temperature and precipitation is largely temporal rather than spatial.

## Model  

Barmah-Millewa is a dynamic ecological system; the habitat suitability for analysed frog species is anticipated to vary according to hydrological conditions; which vary spatially and temporally. While multi-year occupancy models are often analysed using models that estimate local colonisation and extinction rates at each site [@MacKenzie2003], this process becomes more difficult when data is sparse (i.e., only 5 out of 36 sites were measured consistently across the six survey years). When data is sparse and sites are not surveyed consistently then transition parameters (colonisation and extinction) at sites become more difficult to estimate. Given the sparsity of data, and that site characteristics can vary substantially between years we opt for use of a 'stacked' multiseason model. A 'stacked' multiseason model effectively models site-year combinations independently (as different sites), accounting for a particular sites non-independence across years with a site-level random effect [@Kellner2022]. This approach may be particularly well-suited to our data, as primary drivers of occupancy and call activity are temporally variable (hydrological conditions and climate). 

Unlike traditional occupancy models that account for imperfect detection, we modelled occupancy and call activity using a continuous score bounded between 0 and 1 (i.e., the 'probability scores'). To do this we implemented the methods from [@Rhinehart2022]. To do this, occupancy at a songmeter-site for each species and a given survey year was modelled as:  

$$z_{si} \sim Bernoulli(\psi_{si})$$
Whereby $\psi_i$ is the probability of occupancy at that songmeter-site-year; modelled as a logit-link linear model: 

$$\psi_{s,i} = \alpha_{s} + \beta_s \chi_i + \epsilon_{j(s,i)} + \zeta_{y(s,i)} $$
Here, $\alpha_s$ is the intercept for species *s*, $\beta_s \chi_i$ is the contribution of fixed-effect covariates. Additionally random effects for songmeter site (*j*) and year (*y*) are denoted by $\epsilon_{j(s,i)}$ and $\zeta_{y(s,i)}$ respectively.  

The call-rate is modelled through a Bernoulli model:

$$f_{k(s,i)} \sim Bernoulli(z_{s,i} \theta_{k(s,i)})$$
Where $f_{k(s,i)}$ denotes the audio snippet *k* at site *i*. The true occupancy at a site here is denoted by $z_{s,i}$, thus when $z_{s,i} = 0$, the species will not be present in any files. If a species is present at a site ($z_{s,i} = 1$), then the frequency the species call appears in files is given by the parameter estimating call rate ($\theta_{k(s,i)}$). We model this parameter as a logit-link linear model:

$$\theta_{k(s,i)} = \delta_{s} + \eta_s \chi_{k(s,i)} + \xi_{k(s,i)} $$
Here, $\delta_{s}$is the intercept, $\eta_s \chi_{k(s,i)}$ is the contribution of the fixed-effects for daily ovariates that might impact calling rate and $\xi_{k(s,i)}$ is a random-effect for each observation period. 

In @Rhinehart2022, a logit-transformed normal distribution is used for the 'probability scores'. However, here we opt into using a beta distribution for modelling classifier performance and relating the 'probability scores' to observations or non-observations. To model this process we parameterize a beta-distribution model with a random-effects linear model, where the random-effect is the contribution to performance of the acoustic classifier for a given species. This allows us to model the performance of the acoustic classifier differently for each species, however it allows us to use 'share' information on classifier performance between species; which may be particularly informative in cases where a given species has fewer validated detections and estimation of performance is more difficult within the model. Modelling acoustic classifier performance, or how probability scores from the acoustic classifier relate to validated detections and non-detections is given as: 

$$p_{k(s,i)} ∣ f_{k(s,i)} = 0 ∼ Beta(\alpha_{1,s}, \beta_{1,s}) $$
$$p_{k(s,i)} ∣ f_{k(s,i)} = 1 ∼ Beta(\alpha_{2,s}, \beta_{2,s}) $$
With separate models for cases when the validated probability score ($p_{k(s,i)}$) is validated as a non-detection ($f_{k(s,i)} = 0$), and a detection ($f_{k(s,i)} = 1$). The parameters determining the shape of the distribution ($\alpha$ and $\beta$) are modelled using mean ($inv\_logit(\mu + \gamma_s)$) and variance components ($exp(\sigma + \kappa_s)$): 

$$\alpha = inv\_logit(\mu + \gamma_s) \times exp(\sigma + \kappa_s) $$
$$\beta = exp(\sigma + \kappa_s)-inv\_logit(\mu + \gamma_s) \times exp(\sigma + \kappa_s) $$
Here $\mu$ is the intercept for the mean and $\sigma$ is the intercept for the variance. Species-level random effects for the mean and variance models are given with $\gamma_s$ and $\kappa_s$ respectively. 

The likelihood functions for each of the possible combinations of file annotations and whether the site is validated as being occupied can be found within Table 1 of [@Rhinehart2022], and our STAN code presented above.

Models were written using the STAN programming language and run for 250 warmup and 250 sampling iterations across 8 parallel chains. Model convergence and diagnostics were determined using 

# Model analysis 

## Posterior predictive checks  

Posterior predictive checks can be used to compare model predictions to the original data supplied to the model. Here we compare summary statistics for the probability scores supplied to the model. Ideally, predictions from the model are able to align with the supplied data. We compare 25%, 50% and 75% confidence intervals for the 'probability scores'.

```{r PPC, fig.height=24, fig.width=18, message = F, fig.cap = "Posterior predictive checks for the 25%, 50%, and 75% quantiles of backpredicted classifier probability scores against the observed quantiles from validated data"}
q95 <- function(x) quantile(x, 0.95, na.rm = T)
q25 <- function(x) quantile(x, 0.25, na.rm = T)
q50 <- function(x) quantile(x, 0.5, na.rm = T)
q75 <- function(x) quantile(x, 0.75, na.rm = T)
sd_90 <- function (x, na.rm = FALSE) {
  quants <- quantile(x, c(0.05, 0.95), na.rm = T)
  x <- x[x < quants[2] & x > quants[1]]
  sqrt(var(if (is.vector(x) || is.factor(x)) x else as.double(x),
    na.rm = na.rm))
}

species_list_cleaned2 <- species_list_cleaned
names(species_list_cleaned2)[6] <- "Spotted Marsh Frog"

ppc_plots_25 <- list()
ppc_plots_50 <- list()
ppc_plots_75 <- list()
for(i in 1:length(species_list_cleaned)) {
ppc_plots_25[[i]] <- posterior_checks_multispecies(model = fit_beta, 
                              model_data = joined_data, 
                              species_index = i,
                              stat = "q25", 
                              title = paste0("25% quantile ", names(species_list_cleaned2)[i]))

ppc_plots_50[[i]] <- posterior_checks_multispecies(model = fit_beta, 
                              model_data = joined_data, 
                              species_index = i,
                              stat = "q50", 
                              title = paste0("50% quantile ", names(species_list_cleaned2)[i]))

ppc_plots_75[[i]] <- posterior_checks_multispecies(model = fit_beta, 
                              model_data = joined_data, 
                              species_index = i,
                              stat = "q75", 
                              title = paste0("75% quantile ", names(species_list_cleaned2)[i]))

}

ppc_grid <- cowplot::plot_grid(plotlist = c(ppc_plots_25, ppc_plots_50, ppc_plots_75), nrow = 6, byrow = F)

ggsave(plot = ppc_grid, filename = "outputs/plots/ppcs_normal.pdf", width = 6000, height = 8000, units = "px")

ppc_grid
```

## Traceplots 

To check model convergence we can visualise traceplots. Good chain-mixing and convergence is evidenced with traceplots from various chains being overlayed and not diverging.

```{r traceplots, fig.height = 34, fig.width = 30}
params_to_trace <- c("mu_sd", 
                     "sigma_sd", 
                     "beta_psi", 
                     "beta_theta")

traces <- mcmc_trace(fit_beta$draws(params_to_trace))

traces
```

## Call recogniser performance  

Using our model we can visualise how our statistical model is able to delineate validated calls into detections an non-detections. Our model determines the performance of the acoustic classifier by the estimation of two hierarchical parameters for each species. These parameters estimate the 'classifier probability score' ($\mu$) when there is no call present in the snippet ($\mu_{det\ =\ 0}$) and when there is a call present in the snippet ($\mu_{det\ =\ 1}$). At the very least, the classifier should have a higher probability score for detections than non-detections ($\mu_{det\ =\ 1}$ > $\mu_{det\ =\ 0}$).

```{r, fig.height = 8, fig.width = 12, message = FALSE, fig.cap = "Fitted beta-distributions of acoustic classifier performance for scores when a snippet had a detection vs non-detection"}

mu_summary <- fit_beta$draws("mu_pred", format = "df") 
mu_summary_long <- pivot_longer(mu_summary, cols = 1:12)
mu_summary_long$`Validated detection` <- rep(c("Non-detection", "Detection"), length.out = nrow(mu_summary_long))
mu_summary_long$Species <- rep(names(species_list_cleaned), each = 2, length.out = nrow(mu_summary_long))

classifier_plot <- ggplot(data = mu_summary_long) +
  stat_density_ridges(aes(x = value, 
                          y = `Validated detection`, 
                          fill = 0.5 - abs(0.5 - stat(ecdf))), 
                      quantile_lines = TRUE, quantiles = 2, geom = "density_ridges_gradient", calc_ecdf = TRUE) +
  scale_fill_gradient(name = "Posterior density", low = delwp_cols[3], high = delwp_cols[1]) +
  scale_x_continuous(expand = c(0, 0), limits = c(0,1), name = "Classifier probability score") +
  scale_y_discrete(expand = c(0, 0)) +
  coord_cartesian(clip = "off") +
  facet_wrap(~Species) + 
  delwp_theme() +
  theme(panel.spacing = unit(2, "lines"))


ggsave(plot = classifier_plot, filename = "outputs/plots/classifier_plot.pdf", width = 3000, height = 2000, units = "px")

classifier_plot
```

Based on these distributions we see that the classifier is more often than not able to delineate true calls from unlikely calls. However. we also see that these distributions overlap. This means there is a relatively large probability space where a 'classifier probability score' could reasonable both be a non-detection or a detection. 

<!-- # ```{r classifier50} -->
<!-- # mu_sum_50 <- mu_summary_long %>%  -->
<!-- #   group_by(Species, `Validated detection`) %>% -->
<!-- #   summarise(prop_greater = sum(value >= 0.5)/n(),  -->
<!-- #             prop_less = sum(value < 0.5)/n()) -->
<!-- #  -->
<!-- # mu_sum_50 -->
<!-- # ``` -->


## Covariates impacting occupancy  

We investigated whether several hydrological and site covariates impacted site occupancy for each species. Hydrological variables can vary across time (over survey years) and across space (over different sites). Alternatively site type variable just differ between sites.  


```{r FixedEffects, fig.height=16, fig.width = 15, warning = FALSE, message = FALSE, fig.cap="Marginal response curves for the various fixed-effect occupancy parameters used in the model."}
# beta_summaries 
beta_summaries <- fit_beta$summary("beta_psi")

beta_table_summary <- beta_summaries %>% 
  mutate(species = factor(rep(names(species_list_cleaned), times = 7), levels = names(species_list_cleaned)), 
         covariate = rep(c("(Intercept)",
              "Wetness", 
              "Photosynthetic Vegetation", 
              "Water", 
              "Site type = Wetland", 
              "Permenant Water", 
              "Semipermenant Water"), each = length(species_list_cleaned))) %>%
  arrange(species) %>%
  select(species, covariate, mean, sd, q5, q95)

format_cov <- function(sp, b, data = beta_summaries, variable = "beta_psi") {
  fd <- data %>%
    filter(variable == !!paste0(variable, "[", sp, ",",b+1,"]"))
  
  paste0("$\\beta$ = ",
         round(fd$mean, 2), 
        " [90% CI: ", 
        round(fd$q5, 2), 
        " – ",
        round(fd$q95, 2),
        "]")
}

ab_joined_list <- list()
for(j in 1:length(species_list_cleaned)) {
beta_draws <- fit_beta$draws("beta_psi", format = "df")

which_sp <- which(stringr::str_detect(string = colnames(beta_draws),
                                    pattern = paste0("beta_psi\\[", j)))

beta_draws <- beta_draws[,which_sp]

ab_marginal_effects <- list()
ab_plot <- list()

ab_to_use <- psi_form

phi_vars <- c(unlist(stringr::str_remove_all(string = labels(terms(psi_form)), pattern =  "log|scale|sqrt|\\)|\\(")), "Water_permanence")
phi_logs <- c(unlist(stringr::str_detect(string = labels(terms(psi_form)), pattern =  "log\\(")), F)
phi_sqrts <- c(0.5,1,0.5,1,1,1)

phi_labs <- c("Wetness", 
              "Photosynthetic Vegetation", 
              "Water", 
              "Wetland", 
              "Permenant Water", 
              "Semi-permenant Water")

fac <- c(FALSE, FALSE, FALSE, T, T, T)

for(i in 1:3) {
ab_marginal_effects[[i]] <- marginal_effects_cmd(beta_draws, 
                                              param = "beta_psi", species_index = j,
                                              param_number = i+1, log = phi_logs[i],
                                     model_data = site_data[[j]], 
                                     abundance = FALSE,
                                     pwr = phi_sqrts[i],
                                     model_column = phi_vars[i], 
                                     transition = FALSE) %>%
  mutate(species = names(species_list_cleaned)[j])

}
 ab_joined_list[[j]] <- ab_marginal_effects
}

cov_plot <- list()

for(i in 1:3) {

all_me_data <- bind_rows(list(ab_joined_list[[1]][[i]],
                              ab_joined_list[[2]][[i]],
                              ab_joined_list[[3]][[i]],
                              ab_joined_list[[4]][[i]],
                              ab_joined_list[[5]][[i]],
                              ab_joined_list[[6]][[i]])) %>%
  mutate(param = case_when(param == "wet" ~ "Wetness", 
                           param == "pv" ~ "Photosynthetic Vegetation", 
                           param == "water" ~ "Water"), 
         species = case_when(species == "Spotted Marsh Frog (race unknown)" ~ "Spotted Marsh Frog", TRUE ~ species))
  

cov_plot[[i]] <- marginal_effects_plot_cmd_all(all_me_data %>% 
                                            mutate(param = factor(param, levels = phi_labs)), 
                          col = "DarkGreen", 
                          factor = fac[i],
                          ylab = "Occupancy probability") +
   ggplot2::facet_grid(species~param, scales = "free") +
  delwp_theme() +
  theme(strip.text = ggtext::element_markdown()) +
  xlab("Covariate value") +
  delwp_theme()

}

#### Factor effects #### 
site_type <- list()
water_perm <- list()
for(j in 1:length(species_list_cleaned)) {
beta_draws <- fit_beta$draws("beta_psi", format = "df")

which_sp <- which(stringr::str_detect(string = colnames(beta_draws),
                                    pattern = paste0("beta_psi\\[", j)))

beta_draws <- beta_draws[,which_sp]
site_type[[j]] <- data.frame(variable = rep(c("Creekline", 
                                              "Wetland"), each = nrow(beta_draws)), 
                        value = c(plogis(beta_draws[[1]]), plogis(beta_draws[[1]] + beta_draws[[5]])), 
                        param = "Site type", 
                        species = names(species_list_cleaned)[j])

water_perm[[j]] <- tibble(variable = rep(c("Ephemeral creekline", 
                                               "Permenant creekline", 
                                               "Semi-permenant creekline", 
                                               "Ephemeral wetland", 
                                               "Permenant wetland", 
                                               "Semi-permenant wetland"), each = nrow(beta_draws)),
                              `Site type` = rep(c("Creekline", 
                                              "Wetland"), each = 3*nrow(beta_draws)), 
                              `Water permenance` = rep(c("Ephemeral", 
                                              "Permenant", 
                                              "Semi-permenant"), each = nrow(beta_draws), times = 2),
                        value = c(plogis(beta_draws[[1]]), 
                                  plogis(beta_draws[[1]] + beta_draws[[6]]),
                                  plogis(beta_draws[[1]] + beta_draws[[7]]),
                                  plogis(beta_draws[[1]] + beta_draws[[5]]), 
                                  plogis(beta_draws[[1]] + beta_draws[[5]] + beta_draws[[6]]),
                                  plogis(beta_draws[[1]] + beta_draws[[5]] + beta_draws[[7]])), 
                        param = "Site type and water permenance", 
                        species = names(species_list_cleaned)[j])
}

factor_data <- bind_rows(bind_rows(water_perm))

factor_plot <- factor_data %>% 
  mutate(`Water permenance` = factor(`Water permenance`, levels = c("Ephemeral", 
                                              "Semi-permenant", 
                                              "Permenant"))) %>%
      ggplot2::ggplot(aes(x = `Water permenance`, y = value, fill = `Site type`)) +
      ggplot2::geom_violin(draw_quantiles = c(0.05, 0.5, 0.95),
                           alpha = 0.5, scale = "width", trim = TRUE, adjust = 2,
                           position=position_dodge(width=0.5)) +
      ggplot2::facet_grid(species~param, scales = "free") +
      ggplot2::xlab("Water permenance") +
      ggplot2::scale_fill_manual(values = unname(delwp_cols[c(1,3)])) +
      delwp_theme() +
      # ylab("Contribution to occupancy") +
      ggplot2::theme(axis.title.y = element_blank()) 
  

cov_plot_list <- cov_plot
cov_plot_list[[4]] <- factor_plot
cov_plot_list[[2]] <- cov_plot_list[[2]] + ggplot2::theme(axis.title.y = element_blank())
cov_plot_list[[3]] <- cov_plot_list[[3]] + ggplot2::theme(axis.title.y = element_blank())


cov_plot_joined <- cowplot::plot_grid(plotlist = cov_plot_list, nrow = 1, rel_widths = c(0.22,0.21,0.21,0.35))

ggsave(plot = cov_plot_joined, filename = "outputs/plots/occupancy_covariate_effects.pdf", width = 4000, height = 4250, units = "px")

cov_plot_joined
```


## Covariates impacting call activity  

We also model how various covariates impact nightly/daily calling activity. Here we look at nightly precipitation, minimum temperature, current hydrological 'wetness', 'water' and 'photosynthetic vegetation' cover.

```{r FixedEffectsCallRate, fig.height=15, fig.width = 12, warning = FALSE, message = FALSE, fig.cap="Marginal response curves for the various fixed-effect call-rate parameters used in the model."}
# beta_summaries 
theta_summaries <- fit_beta$summary("beta_theta")

theta_table_summary <- theta_summaries %>% 
  mutate(species = factor(rep(names(species_list_cleaned), times = 9), levels = names(species_list_cleaned)), 
         covariate = rep(c("(Intercept)",
                           "Precipitation", 
                           "Minimum Temperature",
                           "Wetness", 
                           "Water",
                           "Photosynthetic Vegetation", 
                           "Regulated", 
                           "Gates On", 
                           "Regulated:any_gates_on"), each = length(species_list_cleaned))) %>%
  arrange(species) %>%
  select(species, covariate, mean, sd, q5, q95)

ab_joined_list <- list()
for(j in 1:length(species_list_cleaned)) {
theta_draws <- fit_beta$draws("beta_theta", format = "df")

which_sp <- which(stringr::str_detect(string = colnames(theta_draws),
                                    pattern = paste0("beta_theta\\[", j)))

beta_draws <- theta_draws[,which_sp]

ab_marginal_effects <- list()
ab_plot <- list()

ab_to_use <- theta_form

phi_vars <- unlist(stringr::str_remove_all(string = labels(terms(ab_to_use)), pattern =  "log|scale|sqrt|\\)|\\("))
phi_logs <- unlist(stringr::str_detect(string = labels(terms(ab_to_use)), pattern =  "log\\("))
phi_sqrts <- c(1,1,1,1,1)

phi_labs <- c("Precipitation", 
              "Minimum Temperature",
              "Wetness", 
              "Water",
              "Photosynthetic Vegetation")

fac <- c(FALSE, FALSE, FALSE, FALSE, FALSE)

for(i in 1:length(phi_labs)) {
ab_marginal_effects[[i]] <- marginal_effects_cmd(theta_draws, 
                                              param = "beta_theta", species_index = j,
                                              param_number = i+1, log = phi_logs[i],
                                     model_data = joined_data[[j]], 
                                     abundance = FALSE,
                                     pwr = phi_sqrts[i],
                                     model_column = phi_vars[i], 
                                     transition = FALSE) %>%
  mutate(species = names(species_list_cleaned)[j])

}
ab_joined_list[[j]] <- bind_rows(ab_marginal_effects)
}

all_theta_data <- bind_rows(ab_joined_list) %>%
  mutate(param = case_when(param == "wet" ~ "Wetness", 
                           param == "pv" ~ "Photosynthetic Vegetation", 
                           param == "water" ~ "Water", 
                           param == "tmin" ~ "Minimum Temperature", 
                           TRUE ~ param), 
         species = case_when(species == "Spotted Marsh Frog (race unknown)" ~ "Spotted Marsh Frog", TRUE ~ species))
  

theta_plot <- marginal_effects_plot_cmd_all(all_theta_data %>% 
                                            mutate(param = factor(param, levels = phi_labs)), 
                          col = "DarkGreen", 
                          factor = FALSE,
                          ylab = "Nightly call activity/detection probability") +
   ggplot2::facet_grid(species~param, scales = "free") +
  delwp_theme() +
  theme(strip.text = ggtext::element_markdown()) +
  xlab("Covariate value")

## factor plot ##
#### Factor effects #### 
regulator_eff <- list()
for(j in 1:length(species_list_cleaned)) {
beta_draws <- fit_beta$draws("beta_theta", format = "df")

which_sp <- which(stringr::str_detect(string = colnames(beta_draws),
                                    pattern = paste0("beta_theta\\[", j)))

beta_draws <- beta_draws[,which_sp]

regulator_eff[[j]] <- tibble(`Active management` = rep(c("Unregulated", 
                                              "Regulated OFF", 
                                              "Regulated ON"), each = nrow(beta_draws)),
                              `Regulated` = rep(c("Unregulated", 
                                              "Regulated", 
                                              "Regulated"), each = nrow(beta_draws)),
                        value = c(plogis(beta_draws[[1]]), 
                                  plogis(beta_draws[[1]] + beta_draws[[7]]),
                                  plogis(beta_draws[[1]] + beta_draws[[7]] + beta_draws[[8]] + beta_draws[[9]])), 
                        param = "Active management", 
                        species = names(species_list_cleaned)[j])
}

regulator_thta_eff <- bind_rows(bind_rows(regulator_eff))

factor_theta_plot <- regulator_thta_eff %>% 
  mutate(`Active management` = factor(`Active management`, levels = c("Unregulated", 
                                              "Regulated OFF", 
                                              "Regulated ON"))) %>%
      ggplot2::ggplot(aes(x = `Active management`, y = value, fill = `Active management`)) +
      ggplot2::geom_violin(draw_quantiles = c(0.05, 0.5, 0.95),
                           alpha = 0.5, scale = "width", trim = T, adjust = 2,
                           position=position_dodge(width=0.5)) +
      ggplot2::facet_wrap(~species, scales = "free", nrow = 2) +
      ggplot2::xlab("Regulated") +
      ggplot2::ylab("Nightly call activity/detection probability") +
      ggplot2::scale_fill_manual(values = c("grey", unname(delwp_cols[c(1,3)]))) +
      delwp_theme() +
      ylab("Nightly call rate") +
      ggplot2::theme(axis.title.y = element_blank()) 

ggsave(plot = theta_plot, filename = "outputs/plots/call_rate_covariate_effects.pdf", width = 4000, height = 4250, units = "px")
ggsave(plot = factor_theta_plot, filename = "outputs/plots/call_rate_regulator.pdf", width = 3500, height = 2600, units = "px")

theta_plot
```

Calling activity can also be impacted by management strategies. Here the impact of active regulation, which allows water to flow through regulators to certain sites can be assessed. With a 2-day lag incorporated, active regulation shows signs of positively impacting call activity:  

```{r factoroccplot, fig.height=8, fig.width = 12, fig.cap="Impact of active regulation on nightly call-rate activity"}
factor_theta_plot
```

## Occupancy rates across years  

Using our model, we estimate how occupancy rates vary for each species, and whether there is variation in these over the years. 

```{r avrates, fig.width=10, fig.height=12, fig.cap= "Estimated occupancy rates accross all songmeter sites per year for each species"}
zhat <- fit_beta$summary("z_hat")
zhat$year <- rep(sort(unique(site_metadata_sf$`Monitoring year`)), times = length(species_list_cleaned))
zhat$species <- rep(names(species_list_cleaned), each = length(unique(site_metadata_sf$`Monitoring year`)))


zhat_draws <- fit_beta$draws("z_hat", format = "matrix") %>%
  as.data.frame() %>%
  pivot_longer(cols = everything()) %>%
  mutate(year = rep(sort(unique(site_metadata_sf$`Monitoring year`)), length.out = nrow(.)),
         species = rep(names(species_list_cleaned), 
                       each = length(unique(site_metadata_sf$`Monitoring year`)),
                       length.out = nrow(.)))



dfas <- list()
for(i in 1:nspecies_cleaned) {
dfas[[i]] <- data.frame(any_seen = stan_data$any_seen[,i], 
                   yearf = stan_data$year_code[,i]) %>%
  group_by(yearf) %>%
  summarise(det = sum(any_seen), 
            tot = n(), 
            rate = det/tot) %>%
  ungroup() %>%
  mutate(year = sort(unique(site_metadata_sf$`Monitoring year`)),
         species = names(species_list_cleaned)[i])
}

dfas_comb <- bind_rows(dfas) %>%
  mutate(Validated = "Proportion of sites with validated occupancy")

zhat_plot <- zhat_draws %>%
  ggplot() +
  geom_violin(aes(x = year, y = value), draw_quantiles = c(0.5), adjust = 4, fill = "grey95", scale = "width") +
  geom_point(aes(x = year, y = rate, fill = Validated), data = dfas_comb, shape = 21, size = 3) +
  facet_wrap(~species, nrow = 3, scales = "fixed") +
  ylab("Proportion of songmeter sites occupied") +
  xlab("Survey year") +
  scale_fill_manual(values = unname(delwp_cols[3])) +
  delwp_theme() +
  theme(legend.position = "top", legend.title = element_text(size = 0), legend.text = element_text(size = 12))

ggsave(plot = zhat_plot, filename = "outputs/plots/zhat_plot.pdf", width = 3500, height = 3000, units = "px")

zhat_plot
```

## Call rates across years  

We can also estimate the average detected calling activities over the years. This estimate is the average across all sites surveyed that year and represents the average nightly detected call rate. Note that this 'call rate' may be a mixture of the acoustic devices ability to detect the call, the calling activity as well as the abundance of frogs at that site. Note that these estimates are conditional upon the site being occupied. Thus an unoccupied site would not have corresponding call rate estimated.  

```{r avcallrate, fig.width=10, fig.height=12, fig.cap= "Estimated call rates accross all songmeter sites per year for each species, conditional upon the site being occupied"}
theta_hat <- fit_beta$summary("theta_hat")
theta_hat$year <- zhat$year
theta_hat$species <- zhat$species

theta_hat_draws <- fit_beta$draws("theta_hat", format = "matrix") %>%
  as.data.frame() %>%
  pivot_longer(cols = everything()) %>%
  mutate(year = rep(sort(unique(site_metadata_sf$`Monitoring year`)), length.out = nrow(.)),
         species = rep(names(species_list_cleaned),
                       each = length(unique(site_metadata_sf$`Monitoring year`)),
                       length.out = nrow(.)))

theta_hat_plot <- theta_hat_draws %>%
  ggplot() +
  geom_violin(aes(x = year, y = value), adjust = 2, draw_quantiles = 0.5, fill = "grey95") +
  facet_wrap(~species, nrow = 3, scales = "fixed") + 
  scale_y_continuous(name = "Average calling activity (nightly call probability)") +
  xlab("Survey year") +
  delwp_theme()

ggsave(plot = theta_hat_plot, filename = "outputs/plots/theta_hat_plot.pdf", width = 3500, height = 3000, units = "px")

theta_hat_plot
```

## Unconditional call activity (relative activity/abundance) 

Given that call rate is dependent upon the site being occupied, the estimated call rate may be better interpreted alongside occupancy in order to measure the relative frog activity or abundance at the site. A score of 1 means that every site is predicted to be occupied and the species is calling every night. This metric might be beneficial in comparing relative activity and/or abundance between species and across years. The values in the plot are also summarised in the subsequent table.    

```{r relAb, fig.width=10, fig.height=12, fig.cap= "Estimated activity accross all songmeter sites per year for each species, unconditional on occupancy (a mixture of both occupancy and calling activity)"}
rel_ab_draws <- fit_beta$draws("rel_ab_mean", format = "matrix") %>%
  as.data.frame() %>%
  pivot_longer(cols = everything()) %>%
  mutate(year = rep(sort(unique(site_metadata_sf$`Monitoring year`)), length.out = nrow(.)),
         species = rep(names(species_list_cleaned), 
                       each = length(unique(site_metadata_sf$`Monitoring year`)),
                       length.out = nrow(.)))

rel_ab_draws_mean <- rel_ab_draws %>%
  group_by(species) %>%
  summarise(Median = median(value), 
            SD = sd(value), 
            `5% CI` = quantile(value, 0.05), 
            `95% CI` = quantile(value, 0.95)) 

rel_ab_plot <- rel_ab_draws %>%
  ggplot() +
  geom_hline(aes(yintercept = Median), linetype = "dashed", colour = "DarkRed", data = rel_ab_draws_mean) +
  geom_violin(aes(x = year, y = value), adjust = 2, draw_quantiles = 0.5, fill = "grey95", alpha = 0.8) +
  ggtext::geom_richtext(aes(y = Median, x = 0, label = round(Median, 2)), data = rel_ab_draws_mean, angle = 90, nudge_x = 0.6) +
  facet_wrap(~species, nrow = 3, scales = "fixed") + 
  scale_y_continuous(name = "Unconditional calling activity") +
  xlab("Survey year") +
  delwp_theme()

ggsave(plot = rel_ab_plot, filename = "outputs/plots/rel_ab_plot.pdf", width = 3500, height = 3000, units = "px")

rel_ab_plot
```

```{r callratetable}
ab_hat <- fit_beta$summary("rel_ab_mean")
ab_hat$year <- rep(sort(unique(site_metadata_sf$`Monitoring year`)), times = length(species_list_cleaned))
ab_hat$species <- rep(names(species_list_cleaned), each = length(unique(site_metadata_sf$`Monitoring year`)))

ab_hat %>%
  select(Year = year, Species = species, Median = median, SD = sd, `5% CI` = q5, `95% CI` = q95) %>%
  bind_rows(rel_ab_draws_mean %>% mutate(Year = "All") %>% rename(Species = species)) %>%
  arrange(Year) %>% 
  kbl(format = "html", digits = 2, row.names = F, caption = "Unconditional nightly call rates across species and years") %>%
  kable_styling(bootstrap_options = "basic") %>%
  collapse_rows(1:2)
```


## Species-richness 

We can calculate species-richness at each site by estimating how many (out of the 6 species modelled) are estimated to occupy each site. 

```{r sprich, fig.width=9, fig.height=6, fig.cap= "Estimated species-richness across sites and over the years. Species richness is grouped according to measured flood scores at that site, in that particular year. The grey points in the background represent the distribution of posterior samples for the average species-richness across all sites from year-to-year"}
sp_r <- fit_beta$summary("species_richness")

sp_r_draws <- t(fit_beta$draws("species_richness", format = "matrix")) %>% as.data.frame()
sp_r_draws$year <- year_code[[1]]
sp_r_draws$Site_f <- site_data[[1]]$Site_f


sp_r_draws_long <- pivot_longer(sp_r_draws, cols = !all_of(c("year", "Site_f"))) %>%
  group_by(name, year) %>%
  summarise(mean = mean(value)) %>%
  ungroup()

sp_r_draws_long_mean <- sp_r_draws_long %>%
  group_by(year) %>%
  summarise(label = paste0("mean: ", round(mean(mean), 2), "\n[90% CI: ", round(quantile(mean, 0.05), 2), ", ", round(quantile(mean, 0.95), 2), "]"), 
            mean = mean(mean))

sp_r$year <- year_code[[1]]
sp_r$FloodScore <- site_data[[1]]$FloodScore

cols_sep <- unlist(lapply(delwp_cols_seq, function(x) x[c(6,10)]))

sp_r_plot <- sp_r %>%
  ggplot() +
  # geom_hline(yintercept = 5.25, linetype = "dashed", colour = "DarkRed") +
  ggbeeswarm::geom_quasirandom(aes(y = mean, x = year), data = sp_r_draws_long, shape = 21, size = 1, fill = "grey70", alpha = 0.25, stroke = 0.1) +
  ggbeeswarm::geom_quasirandom(aes(y = mean, x = year, fill = FloodScore), shape = 21, size = 2) + 
  geom_label(data = sp_r_draws_long_mean, aes(label = label, y = 5.9, x = year), alpha = 0.8, size = 2.5) +
  scale_fill_manual(values = unname(delwp_cols)) +
  labs(x = "Year", y = "Site-level species richness", fill = "Flood score") +
  scale_x_continuous(breaks = 1:6, labels = paste(2018:2023, 2019:2024, sep = " - ")) +
  delwp_theme()

ggsave(plot = sp_r_plot, filename = "outputs/plots/sp_r_plot.pdf", width = 2800, height = 1800, units = "px")

sp_r_plot
```

